---
title: "Data_manipulation_visualization_classroom"
author: "Sukumar"
date: "March 7, 2017"
output: html_document
---

For our next experiment, we use directmarketing file, where we will practise data manipulation and data visualization

```{r Reading data and library, include=FALSE}
library(tidyverse)
dm <- read_csv("E:/tutorials/Jigsaw/R/Data visualization/Data Visualiazaion in R - Pre class Videos/DirectMarketing.csv")
```

```{r}
head(dm)
dim(dm)
```

# Lets say, we want to find out the top 2 salaries of males and females

We will use pipe ( %>% ) operator for this:
Pipe operation works in the same way as we think
Lets break the operation into small manageable parts
* Select the columns gender and salary
* Group the data according to gender
* Get the ranks of the salaries (in descending order preferably)
* Filter it to the top 2 ranks only of each group

So, now lets try it in our piped code:
```{r}
dm %>% select(Gender, Salary) %>% group_by(Gender) %>% filter(min_rank(desc(Salary))<=2) %>% arrange(desc(Salary)) 
```

#Time to introduce window family of functions

These are the cumulative mathematical functions in R
* `cumsum` - cumulative sum 
* `cummax` - cumulative max
* `cummin` - cumulative min
* `cumany` - checks if ANY of the elements of the logical vector is cumulatively true
* `cumall` - checks if ALL of the elements of the logical vector are cumulatively true
* `cume_dist` - calculates the cumulative percent distributions of the data
```{r}
x <- sample(1:20, 7)
cumsum(x)
cummin(x)
cumall(x <= 5)
cume_dist(x)
cumprod(x) #Calculate cumulative product
cumany(x <= 5)
cummax(x)
```

#Finding the top 2% of both groups
```{r}
dm %>% select(Gender, Salary) %>% group_by(Gender) %>% filter(cume_dist(desc(Salary)) <= 0.02) %>% arrange(desc(Salary))
```

#Dividing numeric vector into bins and dividing the corresponding data frame into different parts

```{r}
dm %>% select(Gender, Salary) %>% mutate(Group = ntile(desc(Salary), 10)) %>% group_by(Group) %>% summarise(
  maximum = max(Salary),
  minimum = min(Salary),
  count = n()
  )
```
Along with dividing the `dm` into 10 groups in descending order of salary, the above code gives the count of each group as well as the maximum and minimum of each group